### 2018-05-26 intro.tex
Fix spelling mistakes using spelling checker

\section{Introduction}\label{sec:intro}
Dense linear algebra libraries are the most fundamental software in
scientific and engineering computing domains.
Basic Linear Algebra Subprograms (BLAS) (FIXME) defines a collection
of APIs which act as standard building blocks for dense matrix operations.
BLAS routines are divided into 3 levels,
level-1 for vector-vector operations,
level-2 for matrix-vetor operations,
and level-3 for matrix-matrix operations.
Processor vendors often provide BLAS implementations
highly optimized for their processors,
such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
The HPC community has also contributed several high-quality
BLAS implementations, e.g. ATLAS~\cite{atlas},
GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
and BLIS~\cite{blis,blisport}.

@@

%% GEMM performs a matrix-multiply-accumulate operation.
Generally, an optimized GEMM implementation consists of two components,
(1) a highly optimized kernel routine to accomplish the computation, and
(2) an overall strategy to partition the workload into small tasks
and schedule these tasks to be executed effectively on the target processor.
The kernel routine is a serial program performing matrix-multiply-accumulate
on matrices from a single task.
The overall strategy partitions the workload by tiling the loops,
and determines a task schedule by choosing a specific traversal order
in the loop nest's iteration space.
In a multi-threaded context, different tasks (loop tiles) are assigned to
different threads to parallelize the whole operation.
The main target of GEMM optimization is maximizing the floating-point
operation throughput, measured by FLoating-point Operations Per Second ($flops$).
On one hand, optimization of the serial kernel routine concentrates on
instruction scheduling to improve instruction throughput.
On the other hand, the overall strategy is responsible for optimizing
memory accesses to protect the kernel from long memory latencies.
In this paper, we focus on optimizing the memory access of GEMM.

Currently, two techniques, software prefetching and data packing,
are widely adopted in GEMM implementations to speed up memory accesses.
Software prefetch instructions are utilized
to load data into cache before they are used, so memory accesses will not
cause long latencies and floating-point instructions can execute at peak throughput.
To better exploit the capability of on-chip caches,
matrix blocks the kernel operates on are packed into continous memory buffer
whose footprint is within the cache capacity before the kernel routine is called.
Because GEMM is a computational intensive operation whose
arithmetic complexity is $O(n^3)$ and memory complexity is $O(n^2)$,
the overhead caused by data packing is negligible when
the input matrices are large~\cite{gotogemm}.

@@

Following the trend in recent decades, modern architecture design
is introducing more and more cores on a single processor chip.
To reduce cost of on-chip cache memories and coherence networks,
shared cache may become common in future many-core processors.
And non-LRU replacement policies, for example, psudo-random,
may also be used to further reduce the cache design complexity.
Indeed, some high-performance processors based on ARM technology,
have already adopted this design. (FIXME list the processors)
As a consequence, a better solution to the inter-thread cache conflicts problem
on non-LRU shared caches is meaningful.

In this paper, we present a Shared Cache Partitioning (SCP) method,
to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
The key idea is partition the share cache into physically disjoint sets
and assign different sets to different threads. This can be achieved by
exploiting the memory mapping mechanisim of set-associative caches.

The main contributions of this paper are:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with non-LRU shared caches.
\item We implement SCP in the OpenBLAS library and evaluate it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture.
  Phytium 2000+ has 64 cores, private LRU L1 caches
  and psudo-random shared L2 caches.
  Evaluation results show that GEMM performance improves consistently
  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$.
\end{itemize}

>>>>

\section{Introduction}\label{sec:intro}
Dense linear algebra libraries are the most fundamental software in
scientific and engineering computing domains.
Basic Linear Algebra Subprograms (BLAS) (FIXME) defines a collection
of APIs which act as standard building blocks for dense matrix operations.
BLAS routines are divided into 3 levels,
level-1 for vector-vector operations,
level-2 for matrix-vector operations,
and level-3 for matrix-matrix operations.
Processor vendors often provide BLAS implementations
highly optimized for their processors,
such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
The HPC community has also contributed several high-quality
BLAS implementations, e.g. ATLAS~\cite{atlas},
GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
and BLIS~\cite{blis,blisport}.

@@

%% GEMM performs a matrix-multiply-accumulate operation.
Generally, an optimized GEMM implementation consists of two components,
(1) a highly optimized kernel routine to accomplish the computation, and
(2) an overall strategy to partition the workload into small tasks
and schedule these tasks to be executed effectively on the target processor.
The kernel routine is a serial program performing matrix-multiply-accumulate
on matrices from a single task.
The overall strategy partitions the workload by tiling the loops,
and determines a task schedule by choosing a specific traversal order
in the loop nest's iteration space.
In a multi-threaded context, different tasks (loop tiles) are assigned to
different threads to parallelize the whole operation.
The main target of GEMM optimization is maximizing the floating-point
operation throughput, measured by FLoating-point Operations Per Second ($flops$).
On one hand, optimization of the serial kernel routine concentrates on
instruction scheduling to improve instruction throughput.
On the other hand, the overall strategy is responsible for optimizing
memory accesses to protect the kernel from long memory latency.
In this paper, we focus on optimizing the memory access of GEMM.

Currently, two techniques, software prefetching and data packing,
are widely adopted in GEMM implementations to speed up memory accesses.
Software prefetch instructions are utilized
to load data into cache before they are used, so memory accesses will not
cause long latencies and floating-point instructions can execute at peak throughput.
To better exploit the capability of on-chip caches,
matrix blocks the kernel operates on are packed into continuous memory buffer
whose footprint is within the cache capacity before the kernel routine is called.
Because GEMM is a computational intensive operation whose
arithmetic complexity is $O(n^3)$ and memory complexity is $O(n^2)$,
the overhead caused by data packing is negligible when
the input matrices are large~\cite{gotogemm}.

@@

Following the trend in recent decades, modern architecture design
is introducing more and more cores on a single processor chip.
To reduce cost of on-chip cache memories and coherence networks,
shared cache may become common in future many-core processors.
And non-LRU replacement policies, for example, pseudo-random,
may also be used to further reduce the cache design complexity.
Indeed, some high-performance processors based on ARM technology,
have already adopted this design. (FIXME list the processors)
As a consequence, a better solution to the inter-thread cache conflicts problem
on non-LRU shared caches is meaningful.

In this paper, we present a Shared Cache Partitioning (SCP) method,
to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
The key idea is partition the share cache into physically disjoint sets
and assign different sets to different threads. This can be achieved by
exploiting the memory mapping mechanism of set-associative caches.

The main contributions of this paper are:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with non-LRU shared caches.
\item We implement SCP in the OpenBLAS library and evaluate it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture.
  Phytium 2000+ has 64 cores, private LRU L1 caches
  and pseudo-random shared L2 caches.
  Evaluation results show that GEMM performance improves consistently
  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$.
\end{itemize}

### 2018-05-26 related.tex
Fix spelling mistakes using spelling checker

\section{Releated Work}\label{sec:related}
The idea that all level-3 BLAS operations can be built upon
a high-performance GEMM implemention was first proposed
in~\cite{gemmbased1,gemmbased2}.
The GotoBLAS library~\cite{gotoblas}, and its successor OpenBLAS~\cite{openblas},
are instantiated based on this insight.
So optimizing GEMM has always been the central task in developing
dense linear algebra software.
The GEMM optimization has two aspects.
One is developing fast computation kernels and
the other is choosing a proper overall strategy
to optimize memory access.

@@

There are several approaches for obtaining optimized GEMM kernels,
%% (1) assembly programming, (2) auto-tuning, and (3) directive-based programming,
yielding different tradeoffs between performance and portability.
In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
the kernel component of GEMM are written by domain experts in assembly.
In the case of auto-tuning,
ATLAS~\cite{atlas} adopts the auto-tuning method to
automatically generate kernels with different parameters
in C and find the best-performing one by running them
on the actual computing system.
POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} uses a 
directive-based programming approache.
Given a GEMM kernel in C, POET inserts annotations into
the C code to direct source-to-source compiler transformations.
AUGEM uses a template-based method to match
predefined patterns in the C code and transforms the matched
C code sequence into an optimized assembly code sequence.
Both the auto-tuning and directive-based programming approaches
rely on the compiler to transform kernels in C to machine instructions.
Kernel performance can be improved with compiler techniques,
including SIMD vectorization~\cite{Larsen:2000,Zhou:2016,Zhou:2016b,
  Eichenberger2004,GCCSLP2007},
polyhedral optimization~\cite{Bondhugula2008A,Kong:2013}, 
and loop tiling~\cite{Lam1991,Spampinato:2014,Xue00}.
\textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific
abstractions available in the LLVM compiler infrastructure
and proposes a series of domain-specific yet architecture-independent optimizations
to obtain high-performance kernels in a portable way.

The overall strategy of GEMM mainly concerns about
tiling factors, loop orders and parallelization.
The tiling factors $M_r$, $N_r$, $M_c$, $N_c$ and $K_c$
are essential to GEMM performance.
ATLAS~\cite{atlas} relies on auto-tuning to determine optimal
values for these factors.
Analytic techniques~\cite{analytic1,analytic2,blisanalytic} are
can also be used instead of auto-tuning.
These analytic methods generally take into consideration
the cache capacity and associativity
but assume a LRU replacement policy.
As far as we know, this paper is the first to
address the problem of cache sharing with non-LRU replacement policies,
and SCP serves as the first proposal to handle this problem.
\cite{gotogemm} gives a detailed discussion of different looping orders.
GEMM are usually parallelized only on layer 3 (the M loop),
but all loops along the M and N dimensions are potentially parallelizable.
BLIS\cite{blispar} allows the users to specify a sophisticated configuration
that more than one loops of layer 1, 3, 4 and 5
can be simultaneously parallelized.
This nested parallelization can better fit complex architecture features
like multi-sockets and hyperthreading.

>>>>

\section{Related Work}\label{sec:related}
The idea that all level-3 BLAS operations can be built upon
a high-performance GEMM implementation was first proposed
in~\cite{gemmbased1,gemmbased2}.
The GotoBLAS library~\cite{gotoblas}, and its successor OpenBLAS~\cite{openblas},
are instantiated based on this insight.
So optimizing GEMM has always been the central task in developing
dense linear algebra software.
The GEMM optimization has two aspects.
One is developing fast computation kernels and
the other is choosing a proper overall strategy
to optimize memory access.

@@

There are several approaches for obtaining optimized GEMM kernels,
%% (1) assembly programming, (2) auto-tuning, and (3) directive-based programming,
yielding different tradeoffs between performance and portability.
In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
the kernel component of GEMM are written by domain experts in assembly.
In the case of auto-tuning,
ATLAS~\cite{atlas} adopts the auto-tuning method to
automatically generate kernels with different parameters
in C and find the best-performing one by running them
on the actual computing system.
POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} uses a 
directive-based programming approach.
Given a GEMM kernel in C, POET inserts annotations into
the C code to direct source-to-source compiler transformations.
AUGEM uses a template-based method to match
predefined patterns in the C code and transforms the matched
C code sequence into an optimized assembly code sequence.
Both the auto-tuning and directive-based programming approaches
rely on the compiler to transform kernels in C to machine instructions.
Kernel performance can be improved with compiler techniques,
including SIMD vectorization~\cite{Larsen:2000,Zhou:2016,Zhou:2016b,
  Eichenberger2004,GCCSLP2007},
polyhedral optimization~\cite{Bondhugula2008A,Kong:2013}, 
and loop tiling~\cite{Lam1991,Spampinato:2014,Xue00}.
\textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific
abstractions available in the LLVM compiler infrastructure
and proposes a series of domain-specific yet architecture-independent optimizations
to obtain high-performance kernels in a portable way.

The overall strategy of GEMM mainly concerns about
tiling factors, loop orders and parallelization.
The tiling factors $M_r$, $N_r$, $M_c$, $N_c$ and $K_c$
are essential to GEMM performance.
ATLAS~\cite{atlas} relies on auto-tuning to determine optimal
values for these factors.
Analytic techniques~\cite{analytic1,analytic2,blisanalytic} are
can also be used instead of auto-tuning.
These analytic methods generally take into consideration
the cache capacity and associativity
but assume a LRU replacement policy.
As far as we know, this paper is the first to
address the problem of cache sharing with non-LRU replacement policies,
and SCP serves as the first proposal to handle this problem.
\cite{gotogemm} gives a detailed discussion of different looping orders.
GEMM are usually parallelized only on layer 3 (the M loop),
but all loops along the M and N dimensions are potentially parallelizable.
BLIS\cite{blispar} allows the users to specify a sophisticated configuration
that more than one loops of layer 1, 3, 4 and 5
can be simultaneously parallelized.
This nested parallelization can better fit complex architecture features
like multi-sockets and hyper-threading.

### 2018-05-26 scp.tex
Fix spelling mistakes using spelling checker

Suppose we are implementing a double precision GEMM (DGEMM) on
the 4-core processor, i.e. $es=8B$.
We now determine the tiling factors used in DGEMM following the procedure described
by constraints (\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}).
Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
e.g. $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
It takes careful consideration to determine which one is the best,
and a tuning process may be needed.
After tuning we obtain the best values $M_r = 4$ and $N_r = 8$.
By (\ref{eq:constraints.l1}),
$K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
Here we choose $K_c=256$.
By (\ref{eq:constraints.l2}),
$M_c \le \lfloor (c_{L2}/es/nt_{L2} - 2 N_r K_c )/ K_c \rfloor =
\lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
To leave space for other data structures, as well as the program code,
it is reasonable to shrink $M_c$ to a slightly smaller value,
e.g. $192$, $208$ and $224$ are all proper candidates for $M_c$.
Similar to $M_r$ and $N_r$, a tunning process may be needed to
find the best value for $M_c$.
Here we choose $M_c = 192$.
As there is no L3 cache, (\ref{eq:constraints.l3}) can be ignored
and $N_c$ can be given a large value.
Here we choose $N_c = 1024NT = 4096$.
With tiling factors determined, the sizes of packed matrices
can be calculated. Table.~\ref{tab:msizes} list the sizes of
various packed matrices.

@@

Each $A_2$ is packed into a single continous buffer.
Figure.~\ref{fig:partition.conventional} shows how the
$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
We see that every cache set $s \in [0,2048)$ contains
data from all 4 $A_2$ instances,
potentially leading to inter-thread cache conflicts.
What if the $A_2$ matrices are distributed in the way shown in
Figure.~\ref{fig:partition.segmented}?
$A_2$ of different threads live in strictly disjoint cache sets,
and inter-thread cache conflicts caused by $A_2$ are eliminated completely!

Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
essentially represent two different partitioning styles of the shared cache,
refered to as way-partitioning and set-partitioning, respectively.
To guarantee the data layout in set-partitioning,
$A_2$ cannot be stored in a single continous buffer any more.
Instead, $A_2$ is distributed to 12 continous memory segments, each of size $32KB$.
The distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While the set-partitioning of shared caches avoids inter-thread cache conflicts,
it introduces some complexity to GEMM implementation
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continous ones.

How about $B_2$, the other packed matrix use in GEBP?
Unlike $A_2$, which is thread private,
$B_2$ is shared among all threads and
the packing of $B_2$ is done collabratively by all threads.
We have two choices, privatize $B_2$ and apply the set-partitioning method,
or fall back to conventional way-partitioning for $B_2$.
$B_2$ can be made thread private if every thread makes
its own pack of the whole $B_1$ matrix.
The first choice achieves a full per-thread data isolation
at the expense of redundant packing overhead.
The second choice avoids the extra overhead in both time and space
but risks inter-thread conflicts caused by the shared $B_2$ matrix.
The two choices are evaluated and compared in Section~\ref{sec:evaluation}.

\subsection{Formal Description}\label{subsec:formal}
Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierachy,
The SCP method systematically determines the
the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
Memory layout of other packed matrices on lower layers are
determined by $A_2$ and $B_2$.
We assume that the share matrix $B_2$ is packed
in conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

@@

In Algorithm~\ref{alg:scp.phase1},
the per-thread size for memory buffer $\mathcal{D}_p$ and
$\mathcal{D}_s$ are first computed
(line~\ref{line:size.p}--\ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
and $size_p$ is enlarged to a multile of $align$ (line~\ref{line:align}).
$align$ is initialized with the L1 cache line size $l_1$.
Then on each level $l$ (line~\ref{line:align.for})
with shared non-LRU caches (line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
original $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight in the $wc_l/nt_l$ value is that
each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
in the set-partitioning of shared caches.
At last, memory is allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin}--\ref{line:d.end}).

@@

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} shows the second phase of SCP.
The main part is a recursive procedure $subspace$
(line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierachy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
$subspace$ has three input parameters, $l$ and $idx$ to
identify the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e. a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$ and $subspace$ returns early
(line~\ref{line:shortcut.begin}--\ref{line:shortcut.end}).
If this is not the case, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(line~\ref{line:for.begin}--\ref{line:for.end}).
For each child node $cidx$ on layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branches (line~\ref{line:if}--\ref{line:endif})
show two distinct ways to compute $\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partitioning style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partitioning on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierachy in a depth-first-search order.

>>>>

Suppose we are implementing a double precision GEMM (DGEMM) on
the 4-core processor, i.e. $es=8B$.
We now determine the tiling factors used in DGEMM following the procedure described
by constraints (\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}).
Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
e.g. $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
It takes careful consideration to determine which one is the best,
and a tuning process may be needed.
After tuning we obtain the best values $M_r = 4$ and $N_r = 8$.
By (\ref{eq:constraints.l1}),
$K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
Here we choose $K_c=256$.
By (\ref{eq:constraints.l2}),
$M_c \le \lfloor (c_{L2}/es/nt_{L2} - 2 N_r K_c )/ K_c \rfloor =
\lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
To leave space for other data structures, as well as the program code,
it is reasonable to shrink $M_c$ to a slightly smaller value,
e.g. $192$, $208$ and $224$ are all proper candidates for $M_c$.
Similar to $M_r$ and $N_r$, a tuning process may be needed to
find the best value for $M_c$.
Here we choose $M_c = 192$.
As there is no L3 cache, (\ref{eq:constraints.l3}) can be ignored
and $N_c$ can be given a large value.
Here we choose $N_c = 1024NT = 4096$.
With tiling factors determined, the sizes of packed matrices
can be calculated. Table.~\ref{tab:msizes} list the sizes of
various packed matrices.

@@

Each $A_2$ is packed into a single continuous buffer.
Figure.~\ref{fig:partition.conventional} shows how the
$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
We see that every cache set $s \in [0,2048)$ contains
data from all 4 $A_2$ instances,
potentially leading to inter-thread cache conflicts.
What if the $A_2$ matrices are distributed in the way shown in
Figure.~\ref{fig:partition.segmented}?
$A_2$ of different threads live in strictly disjoint cache sets,
and inter-thread cache conflicts caused by $A_2$ are eliminated completely!

Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
essentially represent two different partitioning styles of the shared cache,
referred to as way-partitioning and set-partitioning, respectively.
To guarantee the data layout in set-partitioning,
$A_2$ cannot be stored in a single continuous buffer any more.
Instead, $A_2$ is distributed to 12 continuous memory segments, each of size $32KB$.
The distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While the set-partitioning of shared caches avoids inter-thread cache conflicts,
it introduces some complexity to GEMM implementation
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continuous ones.

How about $B_2$, the other packed matrix use in GEBP?
Unlike $A_2$, which is thread private,
$B_2$ is shared among all threads and
the packing of $B_2$ is done collaboratively by all threads.
We have two choices, privatize $B_2$ and apply the set-partitioning method,
or fall back to conventional way-partitioning for $B_2$.
$B_2$ can be made thread private if every thread makes
its own pack of the whole $B_1$ matrix.
The first choice achieves a full per-thread data isolation
at the expense of redundant packing overhead.
The second choice avoids the extra overhead in both time and space
but risks inter-thread conflicts caused by the shared $B_2$ matrix.
The two choices are evaluated and compared in Section~\ref{sec:evaluation}.

\subsection{Formal Description}\label{subsec:formal}
Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierarchy,
The SCP method systematically determines the
the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
Memory layout of other packed matrices on lower layers are
determined by $A_2$ and $B_2$.
We assume that the share matrix $B_2$ is packed
in conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

@@

In Algorithm~\ref{alg:scp.phase1},
the per-thread size for memory buffer $\mathcal{D}_p$ and
$\mathcal{D}_s$ are first computed
(line~\ref{line:size.p}--\ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
$align$ is initialized with the L1 cache line size $l_1$.
Then on each level $l$ (line~\ref{line:align.for})
with shared non-LRU caches (line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
original $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight in the $wc_l/nt_l$ value is that
each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
in the set-partitioning of shared caches.
At last, memory is allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin}--\ref{line:d.end}).

@@

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} shows the second phase of SCP.
The main part is a recursive procedure $subspace$
(line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierarchy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
$subspace$ has three input parameters, $l$ and $idx$ to
identify the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e. a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$ and $subspace$ returns early
(line~\ref{line:shortcut.begin}--\ref{line:shortcut.end}).
If this is not the case, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(line~\ref{line:for.begin}--\ref{line:for.end}).
For each child node $cidx$ on layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branches (line~\ref{line:if}--\ref{line:endif})
show two distinct ways to compute $\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partitioning style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partitioning on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierarchy in a depth-first-search order.
